---
title: "Using a Workflow to Model a Domain"
bibliography: []
biblio-style: apalike
link-citations: yes
editor_options: 
  markdown: 
    wrap: 80
---

```{r, include = FALSE}
source(file.path(usethis::proj_get(), "vignettes",  "_common.R"))
path <- tempfile("docs-")
dir.create(path, F, T)
```

## Highlights

-   Create a workflow in R that is readable by a domain expert.
-   Facilitates the transition between concept to implementation of the domain
    model.

## What does `use_workflow` do?

Given the workflow name (order one pizza), the sub-domain name (pizza ordering)
and the number of steps in the workflow (3),

When `use_workflow` is called

```{r, eval = FALSE, echo = TRUE}
use_workflow(name = "order one pizza", domain = "pizza ordering", n_step = 3)
```

```{r, eval = TRUE, echo = FALSE}
withr::with_dir(
    path,
    use_workflow(name = "order-one-pizza", domain = "pizza ordering", n_step = 3)
)
file_path <- list.files(path, "order-one-pizza.R", recursive = TRUE)
```

Then the function:

-   Creates a workflow boilerplate with three steps;
-   Places the file at `r file_path`; and
-   Opens the file (only in interactive mode).

## What is a Step

-   The input to a step should always be an environment.
-   The output of a step should always be an environment.

```{r, eval=FALSE, echo=TRUE}
session <- new.env

step <- function(session){
  
  session$system <- list(date = Sys.Date(), os = Sys.info()["sysname"])

  return(session)
}
```

## How to Move from Conception to Implementation?

In this section we customise the boilerplate code to fit the application design.

### Iteration 0 - Decompose a Process into Steps

Before you call `use_workflow`, think about the process you are about to model.
Essentially, you need to decompose the process into sequential steps that
reflect their relationships and dependencies.

<!-- Steps are formed in the imperative -->

Decomposing a process into steps requires cognitive effort, yet you don't need
to delve into details in this stage.

```{=html}
<div class="alert alert-info" <strong="">
Aim to capure 20% of the steps that are roughly 80% of the process. 
</div>
```
```{=html}
<div class="alert alert-warning" <strong="">
Even the best planning is not so omniscient as to get it right the first time.
</div>
```
```{=html}
<div class="alert alert-danger" <strong="">
Each step may (but don't have to) be dependent on its predecessors steps.<br>
Each step must be independent from its succeeding steps.
</div>
```
In our running example, the process of ordering pizza includes, but not limited
to, three steps:

1.  Issue new order;
2.  Order pizza; and
3.  Commit order.

These three necessary steps for ordering pizza happen sequentially in the order
of their writing. In this case, their order is irreversible, each step is
dependent on its predecessors steps.

We start the implementation process by generating boilerplate code for a
workflow that comprises three steps:

The boilerplate code produced by `use_workflow` looks likes this:

```{r code = readLines(fs::path(path, file_path)), eval=FALSE, echo=TRUE}
```

### Iteration 1 - Rename the Steps

In this iteration we rename the generic names of the steps. Each step name,
should reveal you why the step exists, what it does and how it is used.

<!-- naming convention -->

```{r}
writeLines((
    readLines(fs::path(path, file_path))
    %>% stringr::str_replace_all("step_1", "issue_new_order")
    %>% stringr::str_replace_all("step_2", "order_pizza")
    %>% stringr::str_replace_all("step_3", "commit_order")
), fs::path(path, file_path))
```

```{r code = readLines(fs::path(path, file_path)), eval=FALSE, echo=TRUE}
```

<!-- What makes for a good step name? -->

<!-- If this is your first workflow in the project, strive to create as few elements -->

<!-- as possible, and yet have most of the different element types appear once. For -->

<!-- example, if your domain requires you to fit regression models to five target -->

<!-- variables, that is, you need five model objects, then start by modelling a -->

<!-- single model. Yet, include all the necessary steps to fit the single model. In -->

<!-- this case you may want to have the following steps: -->

### Iteration 2 - ???

<!-- A step is a collection of tasks -->

```{r, message=TRUE}
message("hello")
```
