---
title: "Using a Microservice"
bibliography: []
biblio-style: apalike
link-citations: yes
editor_options: 
  markdown: 
    wrap: 80
---

```{r, include = FALSE}
source(file.path(usethis::proj_get(), "vignettes",  "_common.R"))
```

## What does `use_microservice` do?

Given `entrypoint_name` <microservice> and `endpoint_name` <RESTful>,

When `use_microservice` is called

```{r, eval=FALSE, echo=TRUE}
use_microservice(entrypoint_name = "microservice", endpoint_name = "RESTful")
```

Then the function

Generates the necessary files for a working-out-of-the-box microservice. In
particular, the function:

-   Creates two entrypoints, foreground and background, to start the
    microservice

-   Creates an endpoint with default utility, *healthcheck* and *class*, where

    -   *healthcheck* returns status 200 if the microservice is live, and
    -   *class* returns the R object class received by the microservice

-   Creates a unit-test for the endpoint with supporting helper and setup
    scripts

-   Adds the needed packages names to the project DESCRIPTION file

-   Opens all the files (only in interactive mode)

The boilerplate code produced by add\_value\_object looks likes this:

## Spinning up the Microservice

Running the Microservice is done by one of two scripts:

-   `inst/entrypoints/microservice-foreground.R` runs the Microservice at the
    foreground. This option is appropriate during production. However, running
    the Microservice at this mode means the R session is locked and would not
    process CLI or other scripts commands until the Microservice is stopped.

-   `inst/entrypoints/microservice-background.R`runs the Microservice at the

    background as an RStudio job. This option is appropriate during development.
    Running the Microservice at this mode allows the programmer to interact with
    the Microservice by sending it HTTP requests within the same R session.

## Discovering the Microservice Details

Communicating with the Microservice is done through HTTP requests. This means
you need to know what URL exists, what service they provide, and what input they
require.

Firstly, OpenAPI Specification provides visual documentation for the
Microservice API. If the

Microservice is hosted locally on port 8080, then you can browse OpenAPI at
<http://127.0.0.1:8080/__docs__/.>
