---
title: "Packaging R Projects"
bibliography: [../../inst/REFERENCES.bib]
biblio-style: apalike
link-citations: yes
editor_options: 
  markdown: 
    wrap: 80
---

<!-- One line about what the function does -->

`create_pseudo_package` creates a template project masqueraded as an R package.
That means, we develop the R project in an R package-like structure.

1.  The advantages of developing an R project as an R package:

    -   **Familiarity**. If you are already familiar with R packages, then the
        project's folder structure, components, procedures, and tool set are
        already known to you.
    -   **Maturity**. R packages have been developed for decades. As a result,
        the R community has already spent lots of effort on developing auxiliary
        packages to facilitate R package development. In addition, there are
        books, such as [@wickham2015r], that capture the body of knowledge of
        how to develop R packages. These include both the essentials and good
        practice that has been evolving for many years.

2.  The differences between developing an R project and an R package:

    -   **Function**. R packages are tools while R projects are services.
    -   **Protocols**. All R package has to comply with the ***Writing R
        Extensions*** [@RCoreTeam] protocol. In addition, CRAN packages must
        comply with even a stricter
        [protocol](https://cran.r-project.org/web/packages/policies.html). For
        example, "In principle, packages must pass `R CMD check` without
        warnings or significant notes to be admitted to the main CRAN package
        area". In contrast, R projects don't have to comply with any standard or
        protocol. Instead, it is up to the organisation or programmer, to
        selectively choose which protocols the project should adhere to.
    -   **Purpose**. R packages provide tools for solving problems, while R
        project solves a specific problem, often by importing R packages.

```{r, include = FALSE}
source(file.path(usethis::proj_get(), "vignettes",  "_common.R"))
path <- fs::path_tidy(file.path(tempdir(), "template.project"))
fs::dir_create(path, recurse = TRUE)
```

## Creating a Standardized Project

```{r, echo=TRUE, eval=FALSE}
usethis2::create_pseudo_package(path = tempdir())
```

```{r usethis2-create_project, results='hide'}
wd <- usethis::proj_get() 

usethis:::proj_set_(path) 
usethis::create_package(path, open = FALSE)
usethis::use_testthat()

usethis:::proj_set_(wd)
```

```{r, results='markup'}
z <- capture.output(fs::dir_tree(path))
z <- z[-1]
w <- lapply(
  z, function(x){
    cat(x, "\n")
  }
)
```
