---
title: "Using a Microservice"
bibliography: [../../inst/REFERENCES.bib]
biblio-style: apalike
link-citations: yes
editor_options: 
  markdown: 
    wrap: 80
---

```{r, include = FALSE}
source(file.path(usethis::proj_get(), "vignettes",  "_common.R"))
path <- tempfile("docs-")
dir.create(path, F, T)
```

## What does `use_microservice` do?

**Given** `entrypoint_name` (microservice), `endpoint_name` (RESTful), `host` 
(127.0.0.1) and `port` number (8080),

**When** `use_microservice` is called

```{r, eval=FALSE, echo=TRUE}
use_microservice(
    entrypoint_name = "microservice", 
    endpoint_name = "RESTful",
    host = "127.0.0.1",
    port = 8080             
)
```

```{r, eval=TRUE, echo=FALSE}
withr::with_envvar(list(TESTTHAT = ""), {
    withr::with_dir(
        path, {
            desc::description$new("!new")$write("DESCRIPTION")
            use_microservice(
                entrypoint_name = "microservice", 
                endpoint_name = "RESTful",
                host = "127.0.0.1",
                port = 8080             
            )
        }
    )
})
```

**Then** the function generates the necessary files for a working-out-of-the-box
microservice at <http://127.0.0.1:8080/>. In particular, the function creates or
modifies the following files:

-   `./inst/entrypoints/microservice-foreground.R`

-   `./inst/entrypoints/microservice-background.R`

-   .`/inst/endpoints/RESTful.R`

-   `./tests/testthat/test-endpoint-RESTful.R`

-   `./tests/testthat/helpers-xyz.R`

-   `./tests/testthat/setup-xyz.R`

-   `./DESCRIPTION`

## Spinning up the Microservice

Running the Microservice is done by one of two scripts:

-   `inst/entrypoints/microservice-foreground.R` runs the Microservice at the
foreground. This option is appropriate during production. However, running
the Microservice at this mode means the R session is locked and would not
process CLI or other scripts commands until the Microservice is stopped.

-   `inst/entrypoints/microservice-background.R`runs the Microservice at the

background as an RStudio job. This option is appropriate during development.
Running the Microservice at this mode allows the programmer to interact with the
Microservice by sending it HTTP requests within the same R session.

## Discovering the Microservice Details

Communicating with the Microservice is done through HTTP requests. This means
you need to know what URL exists, what service they provide, and what input they
require.

Firstly, OpenAPI Specification provides visual documentation for the
Microservice API. If the Microservice is hosted locally on port 8080, then you
can browse OpenAPI at <http://127.0.0.1:8080//__docs__/>.

Second source of documentation is the Microservice unit-test, located at
`./tests/testthat/test-endpoint-RESTful.R`. The unit-test fully covers the
available services, demonstrates how to send HTTP requests, and provides
examples of how to post-process the returned response into an R `data.frame`.

## Developing and Testing Endpoints

`use_microservice` automatically generates endpoint unit-test that is useful
both for developing and testing the API. The boilerplate code of the unit-test
looks like:

```{r code = readLines(file.path(path, "tests/testthat/test-endpoint-RESTful.R")), eval=FALSE, echo=TRUE}
```
